---
title: "task4_halloween_candy_analysis"
output: html_document
---

```{r}
# loading in required libraries
library(tidyverse)
library(here)
# reading in and assigning data
halloween_candy <- read_csv(here("clean_data/clean_halloween_candy_2015_to_2017.csv"))
```

1. What is the total number of candy ratings given across the three years. (number of candy ratings, not number of raters. Don’t count missing values)
```{r}
# there was 640451 total ratings excl NA
# using tally to negate the NA
# 115079/755530 are NA's, maybe indicating that users were unable to answer or did not have an answer 
halloween_candy %>% 
  group_by(reaction) %>% 
  tally(!is.na(reaction)) %>% 
  summarise(total_ratings = sum(n))
```

2. What was the average age of people who are going out trick or treating and the average age of people not going trick or treating?
```{r}
# 35 years - could be because the users children are now generally at the age to go trick or treating 
halloween_candy %>% 
  filter(going_out == "Yes") %>% 
  summarise(avg_age_yes = round(mean(age)))

# 39 years - could be because the users childrend are now generally at the age to stop going trick or treating 
halloween_candy %>% 
  filter(going_out == "No") %>% 
  summarise(avg_age_no = round(mean(age)))
```

3. For each of joy, despair and meh, which candy bar received the most of these ratings?
```{r}
# there is no meh in 2015 - this is a consideration
# create table to interrogate 
q3 <- halloween_candy %>% 
  group_by(candy, reaction) %>%
  summarise(number = n(),.groups = "drop") 

# JOY candy - pivot table in ordeJor to have reaction types as names to provide count
# any full sized candy bar
q3 %>%
  pivot_wider(names_from = reaction, values_from = number) %>% 
  select(candy, JOY) %>% 
  slice_max(JOY)

# DESPAIR candy - pivot table in order to have reaction types as names to provide count
# mary janes
q3 %>%
  pivot_wider(names_from = reaction, values_from = number) %>% 
  select(candy, DESPAIR) %>% 
  slice_max(DESPAIR)

# MEH candy - pivot table in order to have reaction types as names to provide count
# lollipops
q3 %>%
  pivot_wider(names_from = reaction, values_from = number) %>% 
  select(candy, MEH) %>% 
  slice_max(MEH)
```

4. How many people rated Starburst as despair?
```{r}
# 1864
q3 %>% 
  pivot_wider(names_from = reaction, values_from = number) %>%
  select(candy, DESPAIR) %>% 
  filter(candy == "starburst")
```


For the next three questions, count despair as -1, joy as +1 and meh as 0.

```{r}
reaction_as_number_candy_combined <- halloween_candy %>% 
 mutate(reaction_as_number = case_when(
      reaction == "JOY"     ~  1,
      reaction == "DESPAIR" ~ -1,
      reaction == "MEH"     ~  0
    ))
```


5. What was the most popular candy bar by this rating system for each gender in the dataset?

```{r}
# full sized candy bar, 1533 for male	
reaction_as_number_candy_combined %>% 
  filter(gender == "Male") %>% 
  group_by(candy) %>% 
  summarise(male_rating = sum(reaction_as_number, na.rm = TRUE)) %>% 
  slice_max(male_rating)

# full sized candy bar, 855 for female	
reaction_as_number_candy_combined %>% 
  filter(gender == "Female") %>% 
  group_by(candy) %>% 
  summarise(female_rating = sum(reaction_as_number, na.rm = TRUE)) %>% 
  slice_max(female_rating)
```


6. What was the most popular candy bar in each year?
```{r}
# any full sized candy bar - 4392
reaction_as_number_candy_combined %>% 
  filter(year == "2015") %>% 
  group_by(candy) %>% 
  summarise(rating_2015 = sum(reaction_as_number, na.rm = TRUE)) %>% 
  slice_max(rating_2015)

# any full sized candy bar - 982
reaction_as_number_candy_combined %>% 
  filter(year == "2016") %>% 
  group_by(candy) %>% 
  summarise(rating_2016 = sum(reaction_as_number, na.rm = TRUE)) %>% 
  slice_max(rating_2016)

# any full sized candy bar - 1497
reaction_as_number_candy_combined %>% 
  filter(year == "2017") %>% 
  group_by(candy) %>% 
  summarise(rating_2017 = sum(reaction_as_number, na.rm = TRUE)) %>% 
  slice_max(rating_2017)
```


7. What was the most popular candy bar by this rating for people in US, Canada, UK and all other countries?

```{r}
# any_full_sized_candy_bar	2122	
reaction_as_number_candy_combined %>% 
  filter(country == "usa") %>% 
  group_by(candy) %>% 
  summarise(rating_us = sum(reaction_as_number, na.rm = TRUE)) %>% 
  slice_max(rating_us)

# any_full_sized_candy_bar	248	
reaction_as_number_candy_combined %>% 
  filter(country == "canada") %>% 
  group_by(candy) %>% 
  summarise(rating_canada = sum(reaction_as_number, na.rm = TRUE)) %>% 
  slice_max(rating_canada)

# any_full_sized_candy_bar	33, lindt_truffle	33, rolos	33, tolberone_something_or_other	33	
reaction_as_number_candy_combined %>% 
  filter(country == "uk") %>% 
  group_by(candy) %>% 
  summarise(rating_uk = sum(reaction_as_number, na.rm = TRUE)) %>% 
  slice_max(rating_uk)

# any_full_sized_candy_bar	4468	
reaction_as_number_candy_combined %>% 
  filter(country == "other") %>% 
  group_by(candy) %>% 
  summarise(rating_other = sum(reaction_as_number, na.rm = TRUE)) %>% 
  slice_max(rating_other)
```

