---
title: "task4_halloween_candy_analysis"
output: html_document
---

```{r}
library(tidyverse)

bbcandy2015 <- janitor::clean_names(
               readxl::read_xlsx("../raw_data/boing-boing-candy-2015.xlsx"))

bbcandy2016 <- janitor::clean_names(
               readxl::read_xlsx("../raw_data/boing-boing-candy-2016.xlsx"))

bbcandy2017 <- janitor::clean_names(
               readxl::read_xlsx("../raw_data/boing-boing-candy-2017.xlsx"))
```

```{r}
# comparing variables of all datasets to find similarites 
janitor::compare_df_cols(bbcandy2017, bbcandy2016, bbcandy2015) %>% 
  view()
```   

```{r}
# 2015 cleaning
# renaming variables 
# adding in year column to clarify when joining
pivot_candy15 <- bbcandy2015 %>%
  select(-timestamp) %>% 
  rename(going_out = are_you_going_actually_going_trick_or_treating_yourself,
         age       = how_old_are_you) %>%   
  add_column(year = 2015, .before = 1) 

# pivotting data to enable join with others, and removing non-candy values
pivot_candy15.1 <- pivot_candy15 %>% 
  pivot_longer(cols = 4:124, names_to = "candy", values_to = "reaction") %>% 
  filter(!candy %in% c(
    "cash_or_other_forms_of_legal_tender",
    "dental_paraphenalia",
    "generic_brand_acetaminophen",
    "glow_sticks",
    "broken_glow_stick",
    "creepy_religious_comics_chick_tracts",
    "hugs_actual_physical_hugs",
    "joy_joy_mit_iodine",
    "lapel_pins",
    "mint_leaves",
    "peterson_brand_sidewalk_chalk",
    "vicodin",
    "please_leave_any_remarks_or_comments_regarding_your_choices",
    "please_list_any_items_not_included_above_that_give_you_joy",
    "please_list_any_items_not_included_above_that_give_you_despair",
    "please_estimate_the_degree_s_of_separation_you_have_from_the_following_celebrities_jk_rowling",
    "please_estimate_the_degree_s_of_separation_you_have_from_the_following_celebrities_jj_abrams",
    "please_estimate_the_degree_s_of_separation_you_have_from_the_following_celebrities_beyonce",
    "please_estimate_the_degree_s_of_separation_you_have_from_the_following_celebrities_bieber",
    "please_estimate_the_degree_s_of_separation_you_have_from_the_following_celebrities_kevin_bacon",
    "please_estimate_the_degree_s_of_separation_you_have_from_the_following_celebrities_francis_bacon_1561_1626",
    "please_estimate_the_degrees_of_separation_you_have_from_the_following_folks_bruce_lee",
    "please_estimate_the_degrees_of_separation_you_have_from_the_following_folks_jk_rowling",
    "please_estimate_the_degrees_of_separation_you_have_from_the_following_folks_malala_yousafzai",
    "please_estimate_the_degrees_of_separation_you_have_from_the_following_folks_thom_yorke",
    "please_estimate_the_degrees_of_separation_you_have_from_the_following_folks_jj_abrams",
    "please_estimate_the_degrees_of_separation_you_have_from_the_following_folks_hillary_clinton",
    "please_estimate_the_degrees_of_separation_you_have_from_the_following_folks_donald_trump",
    "please_estimate_the_degrees_of_separation_you_have_from_the_following_folks_beyonce_knowles",
    "pencils",
    "guess_the_number_of_mints_in_my_hand",
    "betty_or_veronica",
    "check_all_that_apply_i_cried_tears_of_sadness_at_the_end_of",
    "that_dress_that_went_viral_early_this_year_when_i_first_saw_it_it_was",
    "fill_in_the_blank_imitation_is_a_form_of",
    "fill_in_the_blank_taylor_swift_is_a_force_for",
    "what_is_your_favourite_font",
    "if_you_squint_really_hard_the_words_intelligent_design_would_look_like",
    "which_day_do_you_prefer_friday_or_sunday")
      )
```


```{r}
# 2016 cleaning
# renaming variables 
# adding in year column to clarify when joining
pivot_candy16 <- bbcandy2016 %>%
  select(-timestamp) %>% 
  rename(going_out  = are_you_going_actually_going_trick_or_treating_yourself,
         age        = how_old_are_you,
         gender     = your_gender,
         country    = which_country_do_you_live_in,
         province   = which_state_province_county_do_you_live_in) %>%   
  add_column(year = 2016, .before = 1) 

# pivotting data to enable join with others, and removing non-candy values
pivot_candy16.1 <- pivot_candy16 %>% 
  pivot_longer(cols = 7:123, names_to = "candy", values_to = "reaction") %>% 
  filter(!candy %in% c(
    "cash_or_other_forms_of_legal_tender",
    "dental_paraphenalia",
    "generic_brand_acetaminophen",
    "glow_sticks",
    "broken_glow_stick",
    "creepy_religious_comics_chick_tracts",
    "hugs_actual_physical_hugs",
    "joy_joy_mit_iodine",
    "lapel_pins",
    "mint_leaves",
    "peterson_brand_sidewalk_chalk",
    "vicodin",
    "please_leave_any_remarks_or_comments_regarding_your_choices",
    "please_list_any_items_not_included_above_that_give_you_joy",
    "please_list_any_items_not_included_above_that_give_you_despair",
    "please_estimate_the_degree_s_of_separation_you_have_from_the_following_celebrities_jk_rowling",
    "please_estimate_the_degree_s_of_separation_you_have_from_the_following_celebrities_jj_abrams",
    "please_estimate_the_degree_s_of_separation_you_have_from_the_following_celebrities_beyonce",
    "please_estimate_the_degree_s_of_separation_you_have_from_the_following_celebrities_bieber",
    "please_estimate_the_degree_s_of_separation_you_have_from_the_following_celebrities_kevin_bacon",
    "please_estimate_the_degree_s_of_separation_you_have_from_the_following_celebrities_francis_bacon_1561_1626",
    "please_estimate_the_degrees_of_separation_you_have_from_the_following_folks_bruce_lee",
    "please_estimate_the_degrees_of_separation_you_have_from_the_following_folks_jk_rowling",
    "please_estimate_the_degrees_of_separation_you_have_from_the_following_folks_malala_yousafzai",
    "please_estimate_the_degrees_of_separation_you_have_from_the_following_folks_thom_yorke",
    "please_estimate_the_degrees_of_separation_you_have_from_the_following_folks_jj_abrams",
    "please_estimate_the_degrees_of_separation_you_have_from_the_following_folks_hillary_clinton",
    "please_estimate_the_degrees_of_separation_you_have_from_the_following_folks_donald_trump",
    "please_estimate_the_degrees_of_separation_you_have_from_the_following_folks_beyonce_knowles",
    "pencils",
    "guess_the_number_of_mints_in_my_hand",
    "betty_or_veronica",
    "check_all_that_apply_i_cried_tears_of_sadness_at_the_end_of",
    "that_dress_that_went_viral_early_this_year_when_i_first_saw_it_it_was",
    "that_dress_that_went_viral_a_few_years_back_when_i_first_saw_it_it_was",
    "fill_in_the_blank_imitation_is_a_form_of",
    "fill_in_the_blank_taylor_swift_is_a_force_for",
    "what_is_your_favourite_font",
    "if_you_squint_really_hard_the_words_intelligent_design_would_look_like",
    "which_day_do_you_prefer_friday_or_sunday",
      "chardonnay", 
      "person_of_interest_season_3_dvd_box_set_not_including_disc_4_with_hilarious_outtakes",
      "please_leave_any_witty_snarky_or_thoughtful_remarks_or_comments_regarding_your_choices",
      "what_is_your_favourite_font",
      "do_you_eat_apples_the_correct_way_east_to_west_side_to_side_or_do_you_eat_them_like_a_freak_of_nature_south_to_north_bottom_to_top",
      "when_you_see_the_above_image_of_the_4_different_websites_which_one_would_you_most_likely_check_out_please_be_honest",
      "york_peppermint_patties_ignore")
      )
```

```{r}
# cleaning 2017
# This strips out the "q" from the beginning of the column names
names(bbcandy2017) = gsub(pattern = "^q[0-9]*_", replacement = "", 
                           x = names(bbcandy2017))

# removing variables not required for analysis questions
# editing variable names for ease of reference
# adding in year column to clarify when joining
pivot_candy17 <- bbcandy2017 %>%
  select(-c(internal_id, 110:120)) %>% 
  rename(province    = state_province_county_etc,
         mary_janes  = anonymous_brown_globs_that_come_in_black_and_orange_wrappers_a_k_a_mary_janes) %>%   
  add_column(year = 2017, .before = 1) 

# pivotting data to enable join with others, and removing non-candy values
pivot_candy17.1 <- pivot_candy17 %>% 
  pivot_longer(cols = 7:109, names_to = "candy", values_to = "reaction") %>% 
  filter(!candy %in% c(
    "cash_or_other_forms_of_legal_tender",
    "dental_paraphenalia",
    "generic_brand_acetaminophen",
    "glow_sticks",
    "broken_glow_stick",
    "creepy_religious_comics_chick_tracts",
    "hugs_actual_physical_hugs",
    "joy_joy_mit_iodine",
    "lapel_pins",
    "mint_leaves",
    "peterson_brand_sidewalk_chalk",
    "vicodin",
    "please_leave_any_remarks_or_comments_regarding_your_choices",
    "please_list_any_items_not_included_above_that_give_you_joy",
    "please_list_any_items_not_included_above_that_give_you_despair",
    "please_estimate_the_degree_s_of_separation_you_have_from_the_following_celebrities_jk_rowling",
    "please_estimate_the_degree_s_of_separation_you_have_from_the_following_celebrities_jj_abrams",
    "please_estimate_the_degree_s_of_separation_you_have_from_the_following_celebrities_beyonce",
    "please_estimate_the_degree_s_of_separation_you_have_from_the_following_celebrities_bieber",
    "please_estimate_the_degree_s_of_separation_you_have_from_the_following_celebrities_kevin_bacon",
    "please_estimate_the_degree_s_of_separation_you_have_from_the_following_celebrities_francis_bacon_1561_1626",
    "please_estimate_the_degrees_of_separation_you_have_from_the_following_folks_bruce_lee",
    "please_estimate_the_degrees_of_separation_you_have_from_the_following_folks_jk_rowling",
    "please_estimate_the_degrees_of_separation_you_have_from_the_following_folks_malala_yousafzai",
    "please_estimate_the_degrees_of_separation_you_have_from_the_following_folks_thom_yorke",
    "please_estimate_the_degrees_of_separation_you_have_from_the_following_folks_jj_abrams",
    "please_estimate_the_degrees_of_separation_you_have_from_the_following_folks_hillary_clinton",
    "please_estimate_the_degrees_of_separation_you_have_from_the_following_folks_donald_trump",
    "please_estimate_the_degrees_of_separation_you_have_from_the_following_folks_beyonce_knowles",
    "pencils",
    "guess_the_number_of_mints_in_my_hand",
    "betty_or_veronica",
    "check_all_that_apply_i_cried_tears_of_sadness_at_the_end_of",
    "that_dress_that_went_viral_early_this_year_when_i_first_saw_it_it_was",
    "that_dress_that_went_viral_a_few_years_back_when_i_first_saw_it_it_was",
    "fill_in_the_blank_imitation_is_a_form_of",
    "fill_in_the_blank_taylor_swift_is_a_force_for",
    "what_is_your_favourite_font",
    "if_you_squint_really_hard_the_words_intelligent_design_would_look_like",
    "which_day_do_you_prefer_friday_or_sunday",
      "chardonnay", 
      "person_of_interest_season_3_dvd_box_set_not_including_disc_4_with_hilarious_outtakes",
      "please_leave_any_witty_snarky_or_thoughtful_remarks_or_comments_regarding_your_choices",
      "what_is_your_favourite_font",
      "do_you_eat_apples_the_correct_way_east_to_west_side_to_side_or_do_you_eat_them_like_a_freak_of_nature_south_to_north_bottom_to_top",
      "when_you_see_the_above_image_of_the_4_different_websites_which_one_would_you_most_likely_check_out_please_be_honest",
      "york_peppermint_patties_ignore",
         "real_housewives_of_orange_county_season_9_blue_ray")
      )
```

```{r}
# comparing subsets 
janitor::compare_df_cols(pivot_candy15.1, pivot_candy16.1, pivot_candy17.1) %>% 
  view()
```

```{r}
# joining datasets
candy_combined1.0 <- bind_rows(pivot_candy15.1, pivot_candy16.1, pivot_candy17.1)
```

```{r}
# creating pattern for usa and uk country to clean 
usa_pattern <- c("usausausa|usas|usaa|usa? hard to tell anymore..|usa!!!!!!|usa! usa!|usa!|usa usa usa!!!!|usa usa usa usa|usa usa usa|usa (i think but it's an election year so who can really tell)|usa|usa|units states|united ststes|united      ststes|usa|us|usa|united states of america|us|united states|united states|us|u.s.|u.s.a.|united states of america|america|america|u.s.|murica|unites states|us of a|united state|usa! usa! usa!|merica|the united states|united sates|united stated|ussa|'merica|ahem....amerca|n. america|the best one - usa|the united states of america|the yoo ess of aaayyyyyy|u s|u s a|u.s.|u.s.a.|unied states|unhinged states|unied states|unite states|united  states of america|united staes|united statea|united states|united states|united statss|united stetes|	united ststes|united ststes|units states|usa|usa|usa (i think but it's an election year so who can really tell)|usa usa usa|usa usa usa usa|usa usa usa!!!!|usa!|usa! usa!|usa!!!!!!|usa? hard to tell anymore..|usaa|usas|usausausa|murrika|north carolina|california|new jersey|new york|pittsburgh")

uk_pattern <- c("endland|england|scotland|u.k.|uk|united kindom|united kingdom")
```

```{r}
# lowering case of country, making uk, usa, canada and other data only
candy_combined1.2 <-  candy_combined1.0 %>% 
  mutate(country = str_to_lower(country)) %>% 
  mutate(country = case_when(
    str_detect(country, "not[\\s]{1,}") ~ "other",
    str_detect(country, "australia") ~ "other",
    str_detect(country, "austria") ~ "other",
    str_detect(country, "soviet canuckistan") ~ "other",
    str_detect(country, "subscribe to dm4uz3 on youtube") ~ "other",
    str_detect(country, str_c(usa_pattern, collapse = "|")) ~ "usa",
    str_detect(country, str_c(uk_pattern, collapse = "|")) ~ "uk",
    str_detect(country, "^can") ~ "canada",
    is.na(country) == TRUE ~ "other",
    TRUE ~ "other")
  ) 
```

```{r}
# cleaning age column, turning into integer and keeping data from 100 year olds and less
candy_combined1.3 <- candy_combined1.2 %>%
  mutate(age = as.integer(age)) %>% 
  mutate(age = ifelse(age <= 100, age, NA_integer_)) %>% 
  drop_na(age) 
```

1. What is the total number of candy ratings given across the three years. (number of candy ratings, not number of raters. Don’t count missing values)
```{r}
candy_combined1.3 %>% 
  remove_rownames(vicodin, person_of_interest_season_3_dvd_box_set_not_including_disc_4_with_hilarious_outtakes)
```

2. What was the average age of people who are going out trick or treating and the average age of people not going trick or treating?
3. For each of joy, despair and meh, which candy bar received the most of these ratings?
    How many people rated Starburst as despair?
